---
title: "Week 1 project"
author: "Julia"
date: "May 16, 2015"
output: html_document
---
The data for this assignment is downloaded from the course web site.
The variables included in this dataset are:

* steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
* date: The date on which the measurement was taken in YYYY-MM-DD format
* interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Assignment

###Loading and preprocessing the data
```{r}
Activity <- read.csv(file="activity.csv", header=TRUE)
```
###1.What is mean total number of steps taken per day?
```{r as1}
library(dplyr)
Day_Total<-summarise(group_by(Activity, date), sum =sum(steps))
hist(Day_Total$sum)
summary(Day_Total)
avg <-mean(Day_Total$sum, na.rm=TRUE)
med <-median(Day_Total$sum, na.rm=TRUE)
```
The mean of the total number of steps take per day is `r avg` and the median is `r med`.

###2.What is the average daily activity pattern?
```{r as2}
Int_Avg<-summarise(group_by(Activity, interval), avg =mean(steps,na.rm=TRUE))
plot(avg ~ interval,data=Int_Avg, xlab="5 min Intervals", ylab="Average steps across all days", type="l")
summary(Int_Avg)
head(Int_Avg)
max_row <-filter(Int_Avg, Int_Avg$avg==max(Int_Avg$avg))
```
`r max_row$interval`, 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps, `r max_row$avg`.

###3.Imputing missing values
```{r as3}
library(Hmisc)
#Activity$Imp_steps<-impute(Activity$steps, mean)
NA_Steps <-is.na(Activity$steps)
Missing <-length(which(NA_Steps))
Activity$steps[which(is.na(Activity$steps))]<-0

 
IDay_Total<-summarise(group_by(Activity, date), sum =sum(steps))
hist(IDay_Total$sum)
summary(IDay_Total)
Iavg <-mean(IDay_Total$sum, na.rm=TRUE)
Imed <-median(IDay_Total$sum, na.rm=TRUE)
diff_avg<- Iavg-avg
diff_med<- Imed-med
```
The total number of missing values in the dataset is ` r Missing`. After imputing using mean for 5 - minutes interval, the mean of the total number of steps take per day is 'r Iavg` and the median is `r Imed`. After imputing, the mean increased by `r diff_avg` and median increased by `r diff_med`.

###4.Are there differences in activity patterns between weekdays and weekends?
```{r as4}
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
Activity$wDay <-  factor((weekdays(as.Date(Activity$date)) %in% weekdays1)+1L,
      levels=1:2, labels=c('weekend', 'weekday'))
wday_Avg<-summarise(group_by(Activity, interval), wdavg =mean(steps,na.rm=TRUE))
wend_Avg<-summarise(group_by(Activity, interval), wdavg =mean(steps,na.rm=TRUE))
plot(wdavg ~ interval,data=wday_Avg, xlab="5 min Intervals", ylab="Average steps across all days", type="l")

head(Activity)
```
